### 2019-07-23 manuscript/specnet_conffmt.tex
Update keywords.

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[CSS'19]{CSSSA's Annual Conference on Computational Social
Science}{October 24--27, 2019}{Santa Fe, NM, USA}
\acmYear{2019}

\copyrightyear{2019}

\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
\title{Wealth dynamics in the presence of\\network structure and primitive cooperation}
%\titlenote{Produces the permission block, and
 % copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
 % \texttt{acmart.pdf} document}

\author{Rajesh Venkatachalapathy}
\affiliation{%
  \institution{Systems Science Graduate Program, Portland State University}
  \streetaddress{1604 SW 10th Avenue }
  \city{Portland}
  \state{Oregon}
  \postcode{97207-0751}
}
\email{venkatr@pdx.edu}

\author{Stephen Davies}
\affiliation{%
  \institution{University of Mary Washington\\Department of Computer Science}
  \streetaddress{1301 College Avenue}
  \city{Fredericksburg}
  \state{Virginia}
  \postcode{22401}
}
\email{stephen@umw.edu}

\author{William Nehrboss}
\affiliation{%
  \institution{Lake Anna Homeschool}
  \streetaddress{2663 Moody Town RD}
  \city{Bumpass}
  \state{Virginia}
  \postcode{23024}
}
\email{Wknehrboss@gmail.com}

% The default list of authors is too long for headers.
%\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
We study wealth accumulation dynamics in a population of heterogeneously mixed agents with a capacity for a certain primitive form of cooperation enabled by static network structures. Despite their simplicity, the stochastic dynamics generate inequalities in wealth reminiscent of real-world social systems even in a fully mixed population. A simple form of cooperation is introduced and is shown to enhance the viability of agents by embedding such dynamics in a network; the impact of social structures on the origins and persistence of inequality can be teased out easily. The models developed here complement traditional modeling approaches based on grid worlds.

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010147.10010341</concept_id>
<concept_desc>Computing methodologies~Modeling and simulation</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010341.10010349.10010355</concept_id>
<concept_desc>Computing methodologies~Agent / discrete models</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Modeling and simulation}
\ccsdesc[500]{Computing methodologies~Agent / discrete models}

\keywords{economic inequality, dynamical systems, social networks}

\maketitle

\input{bodycssa19}

\bibliographystyle{ACM-Reference-Format}
%\bibliography{sample-bibliography_test}
\bibliography{ref_cssa19}
\end{document}