### 2019-07-23 manuscript/specnet_conffmt.tex
minor changes, Will's comments addressed

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[CSS'19]{CSSSA's Annual Conference on Computational Social
Science}{October 24--27, 2019}{Santa Fe, NM, USA}
\acmYear{2019}

\copyrightyear{2019}

\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
\title[Wealth dynamics in the presence of network structure and primitive cooperation]{Wealth dynamics in the presence of\\network structure and primitive cooperation}
%\titlenote{Produces the permission block, and
 % copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
 % \texttt{acmart.pdf} document}

\author{Rajesh Venkatachalapathy}
\affiliation{%
  \institution{Systems Science Graduate Program, Portland State University}
  \streetaddress{1604 SW 10th Avenue }
  \city{Portland}
  \state{Oregon}
  \postcode{97207-0751}
}
\email{venkatr@pdx.edu}

\author{Stephen Davies}
\affiliation{%
  \institution{University of Mary Washington\\Department of Computer Science}
  \streetaddress{1301 College Avenue}
  \city{Fredericksburg}
  \state{Virginia}
  \postcode{22401}
}
\email{stephen@umw.edu}

\author{William Nehrboss}
\affiliation{%
  \institution{Lake Anna Homeschool}
  \streetaddress{2663 Moody Town RD}
  \city{Bumpass}
  \state{Virginia}
  \postcode{23024}
}
\email{Wknehrboss@gmail.com}

% The default list of authors is too long for headers.
%\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
We study wealth accumulation dynamics in a population of heterogeneously mixed agents with a capacity for a certain primitive form of cooperation enabled by static network structures. Despite their simplicity, the stochastic dynamics generate inequalities in wealth reminiscent of real-world social systems even in a population without social structure. A simple form of social structure and cooperation is introduced and is shown to enhance the viability of agents. The interaction between social structures and dynamics illuminates their role in the generation and persistence of inequality. The models developed here complement traditional modeling approaches based on grid worlds.

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010147.10010341</concept_id>
<concept_desc>Computing methodologies~Modeling and simulation</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010341.10010349.10010355</concept_id>
<concept_desc>Computing methodologies~Agent / discrete models</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Modeling and simulation}
\ccsdesc[500]{Computing methodologies~Agent / discrete models}

\keywords{economic inequality, dynamical systems, social networks}

\maketitle

\input{bodycssa19}

\bibliographystyle{ACM-Reference-Format}
%\bibliography{sample-bibliography_test}
\bibliography{ref_cssa19}
\end{document}